Reproducibility of published meta-analyses on clinical psychological interventions

### 2023-04-21 response_letter.Rmd

- P. 16, line 6: you describe that some discrepancies could be explained by rounding issues. Could you explain this in a bit more detail? Does this mean that there was incorrect rounding in the meta-analyses? And did you count that as correct because it was only a small discrepancy? If so, why? Similarly for the minor reporting errors you mention in line 7: it sounds as if minor errors are not considered issues for the reproducibility. This is a choice that needs to be justified.}

We agree that this could have been more clearly specified. In order to improve our explanation, we think that the clearest way of providing this information is by reporting each of the 15 cases. In this regard, we have added a table in the supplementary file which provides an explanation about each of these 15 cases (see Supplementary file, Qualitative assessment section).

We have added a footnote on p.15 linking to the supplementary file:

>Furthermore, 15 were labelled as \textcolor{blue}{approximately reproduced or reproduced with minor adjustment} in a qualitative check because the discrepancy was probably explained by rounding issues, inverted signs for results (when effect sizes were reported in absolute values) and primary data, minor reporting errors, or minor adjustments in the analytical scheme^[Full details in Supplementary File at https://osf.io/fjhpw].

### 2023-04-21 submission.Rmd
Revised version

```
Of these, `r nrow(df_to_reanalyse)` were re-analysed again as some coding errors were found in the second stage. After this, `r nrow(filter(df_to_reanalyse, numerical_error != "error"))` were re-labelled as reproduced and `r nrow(filter(df_to_reanalyse, numerical_error == "error"))` still had relevant discrepancies. Furthermore, `r nrow(filter(df_mas_revised, qualitative_check == "rep"))` were labelled as \textcolor{blue}{approximately reproduced or reproduced with minor adjustment} in a qualitative check because the discrepancy was probably explained by rounding issues, inverted signs for results (when effect sizes were reported in absolute values) and primary data, minor reporting errors, or minor adjustments in the analytical scheme^[Full details in Supplementary File]. In the remaining `r nrow(filter(df_mas_revised, qualitative_check == "request_information"))`, and in the `r nrow(filter(df_to_reanalyse, numerical_error == "error"))` re-analysed again without success, some issues or relevant discrepancies without apparent explanation were found. 
Figure 5 displays a scatterplot showing the consistency between the original and reproduced summary effect size and their confidence bounds of these `r nrow(df_to_scatter)` meta-analyses. 
Additionally, as a secondary analysis, the reproducibility of the $I^2$ heterogeneity statistic was explored. Figure 6 displays a scatterplot showing the consistency between the original and reproduced $I^2$ statistics. As shown in Figures 5 and 6, the discrepancies found in the heterogeneity statistic $I^2$ are larger than those found in the summary effects and their confidence intervals. \textcolor{blue}{The Pearson's correlation between the summary effect and $I^2$ discrepancies was .172.} The lack of precision of the available data (rounded data) or incomplete information on aspects such as the tau-squared estimator applied seem to have a substantial impact on the reproducibility of this result.

>>>>

```
Of these, `r nrow(df_to_reanalyse)` were re-analysed again as some coding errors were found in the second stage. After this, `r nrow(filter(df_to_reanalyse, numerical_error != "error"))` were re-labelled as reproduced and `r nrow(filter(df_to_reanalyse, numerical_error == "error"))` still had relevant discrepancies. Furthermore, `r nrow(filter(df_mas_revised, qualitative_check == "rep"))` were labelled as \textcolor{blue}{approximately reproduced or reproduced with minor adjustment} in a qualitative check because the discrepancy was probably explained by rounding issues, inverted signs for results (when effect sizes were reported in absolute values) and primary data, minor reporting errors, or minor adjustments in the analytical scheme^[Full details in Supplementary File at: https://osf.io/fjhpw]. In the remaining `r nrow(filter(df_mas_revised, qualitative_check == "request_information"))`, and in the `r nrow(filter(df_to_reanalyse, numerical_error == "error"))` re-analysed again without success, some issues or relevant discrepancies without apparent explanation were found. 
Figure 5 displays a scatterplot showing the consistency between the original and reproduced summary effect size and their confidence bounds of these `r nrow(df_to_scatter)` meta-analyses. 
Additionally, as a secondary analysis, the reproducibility of the $I^2$ heterogeneity statistic was explored. Figure 6 displays a scatterplot showing the consistency between the original and reproduced $I^2$ statistics. As shown in Figures 5 and 6, the discrepancies found in the heterogeneity statistic $I^2$ are larger than those found in the summary effects and their confidence intervals. \textcolor{blue}{The Pearson's correlation between the summary effect and $I^2$ discrepancies was .172.} The lack of precision of the available data (rounded data) or incomplete information on aspects such as the tau-squared estimator applied seem to have a substantial impact on the reproducibility of this result.

### 2023-04-21 submission.tex
Revised version

\abstract{%
Meta-analysis is one of the most useful and powerful research approaches, the relevance of which relies on its credibility. However, in recent years different concerns about the credibility of psychological research have emerged. Reproducibility of scientific results could be considered as the minimal threshold of it. In this study, our purpose was to assess the reproducibility of a sample of published meta-analyses. From a random sample of 100 papers containing at least one meta-analysis on the effectiveness of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files (pdf, doc), or requesting original data from authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis using these data. The original data were retrieved for 146 \textcolor{blue}{(67\%, 146/217)} meta-analyses. Of these, 52 showed a discrepancy larger than 5\% in the main results in the first stage. \textcolor{blue}{For 10 meta-analyses this discrepancy was solved after fixing a coding error and for 15 of them it was solved with minor adjustment or considered approximately reproduced in a qualitative assessment.} In the remaining 27 meta-analyses\textcolor{blue}{(18\%, 27/146)}, different issues were identified in an in-depth review of the papers, such as reporting inconsistencies, lack of data, or transcription errors. \textcolor{blue}{Nevertheless, the numerical discrepancies produced by these issues were mostly minor, with little or no impact on the conclusions.} Current practices of data sharing in meta-analyses hamper the reusability of meta-analytic data. The implementation of new tools would help to avoid certain errors in the meta-analysis reporting process.
}

@@

Of these, 17 were re-analysed again as some coding errors were found in the second stage. After this, 10 were re-labelled as reproduced and 7 still had relevant discrepancies. Furthermore, 15 were labelled as \textcolor{blue}{approximately reproduced or reproduced with minor adjustment} in a qualitative check because the discrepancy was probably explained by rounding issues, inverted signs for results (when effect sizes were reported in absolute values) and primary data, minor reporting errors, or minor adjustments in the analytical scheme\footnote{Full details in Supplementary File}. In the remaining 20, and in the 7 re-analysed again without success, some issues or relevant discrepancies without apparent explanation were found.
Figure 5 displays a scatterplot showing the consistency between the original and reproduced summary effect size and their confidence bounds of these 52 meta-analyses.
Additionally, as a secondary analysis, the reproducibility of the \(I^2\) heterogeneity statistic was explored. Figure 6 displays a scatterplot showing the consistency between the original and reproduced \(I^2\) statistics. As shown in Figures 5 and 6, the discrepancies found in the heterogeneity statistic \(I^2\) are larger than those found in the summary effects and their confidence intervals. \textcolor{blue}{The Pearson's correlation between the summary effect and $I^2$ discrepancies was .172.} The lack of precision of the available data (rounded data) or incomplete information on aspects such as the tau-squared estimator applied seem to have a substantial impact on the reproducibility of this result.

>>>>

\abstract{%
Meta-analysis is one of the most useful and powerful research approaches, the relevance of which relies on its credibility. However, in recent years different concerns about the credibility of psychological research have emerged. Reproducibility of scientific results could be considered as the minimal threshold of it. In this study, our purpose was to assess the reproducibility of a sample of published meta-analyses. From a random sample of 100 papers containing at least one meta-analysis on the effectiveness of interventions in clinical psychology, 217 meta-analyses were selected. We first tried to retrieve the original data by recovering a data file, recoding the data from document files (pdf, doc), or requesting it from original authors. Second, through a multi-stage workflow, we tried to reproduce the main results of each meta-analysis using these data. The original data were retrieved for 146 \textcolor{blue}{(67\%, 146/217)} meta-analyses. Of these, 52 showed a discrepancy larger than 5\% in the main results in the first stage. \textcolor{blue}{For 10 meta-analyses this discrepancy was solved after fixing a coding error and for 15 of them it was solved with minor adjustment or considered approximately reproduced in a qualitative assessment.} In the remaining 27 meta-analyses\textcolor{blue}{(18\%, 27/146)}, different issues were identified in an in-depth review of the papers, such as reporting inconsistencies, lack of data, or transcription errors. \textcolor{blue}{Nevertheless, the numerical discrepancies produced by these issues were mostly minor, with little or no impact on the conclusions.} Current practices of data sharing in meta-analyses hamper the reusability of meta-analytic data. The implementation of new tools would help to avoid certain errors in the meta-analysis reporting process.
}

@@

Of these, 17 were re-analysed again as some coding errors were found in the second stage. After this, 10 were re-labelled as reproduced and 7 still had relevant discrepancies. Furthermore, 15 were labelled as \textcolor{blue}{approximately reproduced or reproduced with minor adjustment} in a qualitative check because the discrepancy was probably explained by rounding issues, inverted signs for results (when effect sizes were reported in absolute values) and primary data, minor reporting errors, or minor adjustments in the analytical scheme\footnote{Full details in Supplementary File at: \url{https://osf.io/fjhpw}}. In the remaining 20, and in the 7 re-analysed again without success, some issues or relevant discrepancies without apparent explanation were found.
Figure 5 displays a scatterplot showing the consistency between the original and reproduced summary effect size and their confidence bounds of these 52 meta-analyses.
Additionally, as a secondary analysis, the reproducibility of the \(I^2\) heterogeneity statistic was explored. Figure 6 displays a scatterplot showing the consistency between the original and reproduced \(I^2\) statistics. As shown in Figures 5 and 6, the discrepancies found in the heterogeneity statistic \(I^2\) are larger than those found in the summary effects and their confidence intervals. \textcolor{blue}{The Pearson's correlation between the summary effect and $I^2$ discrepancies was .172.} The lack of precision of the available data (rounded data) or incomplete information on aspects such as the tau-squared estimator applied seem to have a substantial impact on the reproducibility of this result.
